Report on Late Winter2013 Production: Image Differencing

### 2013-03-27 W13report.tex
Mostly added figures.  Added some text on the fp rate vs. noise estimation

\subsection{Theoretical {\bf ANDYC}}

@@

Section \ref{sec-analyticfp} discusses the theoretical foundation 
that sets the floor on the number of false positives we expect from a 
pure Gaussian random field.  The number of false positives is a strong
function of the seeing and and the confidence level of the detection.  
By comparing the empirical false positive rate to the analytic predictions
we can quantify how close we are getting to the theoretical floor in these
data.  Because the number of false positives is such a strong function of
detection threshold, misestimation of the signal to noise of the detections
has a large impact on how many false positives are observed.  The Figures
\ref{fig-fpthresh1}, \ref{fig-fpthresh2}, and \ref{fig-fpthresh3} show the 
effect of misestimating the noise in the detections for each of three different
seeing values (0.6\", 0.88\", 1.2\" respectively).

To estimate the numbers of false detections we would expect in a
$4000\times4072$ pixel science image with a given FWHM in pixels, we
do the following.  The theoretical expectations for the number of
false detections due to random Gaussian fluctuations is outlined in
\cite{Kaiser-PointSources}.

@@

\subsection{Detection and Measurement on Negative Sources}
We see more negative false positives than positive false positives.  This drove an investigation 
into the symmetry of detection on positive/negative sources.  To do this we looked at the results 
of the image difference task and compared it to the same runs when the difference image was inverted
$differenceImage *= -1$ before measurement.  The results are as follows:
\begin{itemize}
\item The DIA source lists are always the same length.
\item Negative sources always have integer centroid pixel values where positive sources have fractional
pixel precision.
\item Some sources are have the edge bit set in one image or the other.  This leads to a difference in 
the number of false positives as we only consider those without the {\tt EDGE} flag set.
\end{itemize}

@@

>>>>

\subsection{Theoretical {\bf ANDYC}\label{sec-analyticfp}}

@@

Section \ref{sec-analyticfp} discusses the theoretical foundation 
that sets the floor on the number of false positives we expect from a 
pure Gaussian random field.  The number of false positives is a strong
function of the seeing and and the confidence level of the detection.  
By comparing the empirical false positive rate to the analytic predictions
we can quantify how close we are getting to the theoretical floor in these
data.  Because the number of false positives is such a strong function of
detection threshold, misestimation of the signal to noise of the detections
has a large impact on how many false positives are observed.

The Figure \ref{fig-fpthresh} shows the 
effect of misestimating the noise in the detections for each of three different
seeing values (0.6\", 0.88\", 1.2\" respectively).  In each case the top panel shows
the number additional false positives in an LSST size CCD given a fractional misestimation
of the noise in the image and a detection threshold.  A 1\% under-estimate of the noise
corresponds to ~4 additional false positives at a threshold of $5\sigma$ in the best seeing case.  
The bottom panel shows 
the effect of over-estimating the noise in the measurement.  The effect is to reduce
the number of false positves.  A 1\% under-estimate of the noise results in 2 fewer false positives
per chip.

By comparing false positive rates to expected rates from the analytic expressions, we may be able to
estimate the inherent over/under estimation of the reported variance in the images.

%To estimate the numbers of false detections we would expect in a
%$4000\times4072$ pixel science image with a given FWHM in pixels, we
%do the following.  The theoretical expectations for the number of
%false detections due to random Gaussian fluctuations is outlined in
%\cite{Kaiser-PointSources}.

@@

\subsection{Detection and Measurement on Negative Sources}
We see more negative false positives than positive false positives.  This drove an investigation 
into the symmetry of detection on positive/negative sources.  To do this we looked at the results 
of the image difference task and compared it to the same runs when the difference image was inverted
$differenceImage *= -1$ before measurement.  The results are as follows:
\begin{itemize}
\item The DIA source lists are always the same length.
\item Centroiding on negative sources fails almost all the time: 99.7\% of the time sources with negative flux
have the {\tt badcentroid} flag set.
\item psfFlux can be measured at 0. even when no flags are set.
\item psfFlux can be measured as nan with valid centroid values and no other flags set.
\item psfFlux can have the same sign in both the initial and inverted image 0.1\% of measurements.
\item Some sources are have the edge bit set in one image or the other.  This leads to a difference in 
the number of false positives as we only consider those without the {\tt EDGE} flag set.
\end{itemize}

@@

\begin{figure}
\centering
\subfigure[0.6$^{\prime\prime}$ FWHM]{\includegraphics[width=0.45\textwidth]{figures/dfp_v_sigma0.60.eps}}\qquad
\subfigure[0.88$^{\prime\prime}$ FWHM]{\includegraphics[width=0.45\textwidth]{figures/dfp_v_sigma0.88.eps}}\\
\subfigure[1.2$^{\prime\prime}$ FWHM]{\includegraphics[width=0.45\textwidth]{figures/dfp_v_sigma1.20.eps}}
\caption{
We plot here the effect on the number of false positives as a function of detection threshold and noise misestimate.  The three panels 
are for three values of seeing.  In each panel the top pane shows what happens when the noise is under-estimated and the
bottom shows what happens to the false positive rate when the noise is over-estimated.  In all cases the counts are for
the total number of false postives in a single 4000x4072 pixel random Gaussian field.
}
\label{fig-fpthresh}
\end{figure}

\begin{figure}
\includegraphics[width=0.8\textwidth]{figures/fp_v_dsky.eps}
\caption{
Misestimation of the sky value has an effect on both the number of false positives and relative abundance of positive false positives to negative false postives.
The top pane of this figure shows the relative number of +ve/-ve (-ve/+ve) if the sky background is under (over) subtracted at a particular fractional level.  The 
bottom panel shows how the total number of false positves changes for the same misestimation.  This is not a strong function of the seeing, so only the best
seeing example is shown.
}
\label{fig-skythresh}
\end{figure}
